R Notebook

### 2019-03-13 analysis/INLA/test-INLA-sensor.Rmd
fix formula for inla function

```{r}
testdat <- tidy_survey_sets(raw, c("SYN HS","SYN QCS","SYN WCVI","SYN WCHG"), years = c(2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018))
#testdat <- gfplot:::interp_survey_bathymetry(testdat)
glimpse(testdat)
#View(testdat[[1]])
```

```{r}
# get_ssids()
# sensor_syn <- get_sensor_data_trawl(ssid = c(1, 3, 4, 16), spread_attributes = TRUE)
# View(sensor_syn)
# saveRDS(sensor_syn, "data-cache/allsynoptic_sensor_07mar19_time2.rds")
# write.csv(sensor, "synoptic_3_sensor_data_07mar19.csv")
# sensor <- read_csv("data-cache/synoptic_3_sensor_data_07mar19.csv", col_types = cols(do_mlpL_avg = col_double(), 
#                     do_mlpL_max = col_double(), do_mlpL_min = col_double(), salinity_PSU_avg = col_double(), salinity_PSU_max = col_double(), 
#                     salinity_PSU_min = col_double()), na = "NA")
sensor <- readRDS("data-cache/allsynoptic_sensor_07mar19_time2.rds")
# View(sensor)
```

```{r}
dat <- left_join(testdat[[1]], sensor, by="fishing_event_id", suffix = c("", ".sensor"))
# dat <- left_join(testdat, sensor, by="fishing_event_id", suffix = c("", ".sensor"))
dat$year_f <- factor(dat$year)
# View(dat)
dat <- dat %>% mutate(do_range = do_mlpL_max - do_mlpL_min, salinity_range = salinity_PSU_max -
  salinity_PSU_min, temp_range = temperature_C_max - temperature_C_min, depth_range = depth_m_max - depth_m_min)
# glimpse(dat)
```

@@

```{r}
scaled <- scale_survey_predictors(dat)
# glimpse(scaled)
temp_dat <- scaled %>% filter (density!="NA") %>% filter ( ssid =="1") %>% filter(temperature_C_avg!="NA") # %>% filter(year==2017)
# glimpse(temp_dat)
# do_dat <- scaled %>% filter (density!="NA") %>% filter ( ssid =="1") %>%filter(do_mlpL_avg!="NA")
#pg <- readRDS("prediction-grids/QCSpred-grid-interp-cell-width-2.rds")
#pg <- as.array(pg)

@@

```{r}
ggplot(sens_dat, aes(depth)) + geom_histogram() 
ggplot(sens_dat, aes(density*1000)) + geom_histogram() 
ggplot(sens_dat, aes(depth, (density*1000))) + geom_point() 
#ggplot(synhs, aes(depth_m, log(density*1000+1))) + geom_point() 
```

@@

d <- temp_dat
response = "temperature_C_avg"
n_knots = 50
family = "gaussian"
plot = TRUE
fit_model = TRUE
extend = list(n = 8, offset = -0.1)
RangeGuess = 100 
offset = c(RangeGuess/5, RangeGuess)
max_edge = c(RangeGuess/5,RangeGuess)
cutoff = (RangeGuess/5)/5
prior_range = c(RangeGuess, 0.01) # P(range < 0.05) = 0.01
prior_sigma = c(0.5, 0.5)
kmeans = FALSE    
verbose = FALSE
debug = FALSE
trials = 2
covariates = c("depth_scaled", "depth_scaled2")
time = "year"

fit_inla_sensor <- function(d, response = "temperature_C_avg",
                     n_knots = 50,
                     family = "binomial",
                     plot = FALSE,
                     fit_model = TRUE,
                     extend = list(n = 8, offset = -0.1),
                     kmeans = FALSE,                   
                     RangeGuess = 100, 
                     offset = c(RangeGuess/5, RangeGuess),                 
                     max_edge = c(RangeGuess/5,RangeGuess),
                     cutoff = (RangeGuess/5)/5,
                     prior_range = c(RangeGuess, 0.01), # P(range < 0.05) = 0.01
                     prior_sigma = c(0.5, 0.5),
                     time = "year",
                     covariates = NULL,
                     verbose = FALSE,
                     debug = FALSE,
                     trials = 10) {

   # create data.frame including response ("y"), coordinates("xcoo", "ycoo"), covariates, and "intercept"
  dat <- data.frame(
    y = d[, response, drop = TRUE],
    xcoo = d$X,
    ycoo = d$Y, 
    raw_time = d[, time, drop = TRUE],
    select(d, covariates)
  )

  # make a design matrix where the first year is the intercept:
  YO  <- order(dat$raw_time)
  dat <- dat[YO,]

  time_k <- as.numeric(as.factor(dat$raw_time))

  #time_k <- dat$raw_time - min(dat$raw_time) + 1
  k <- max(time_k)

  time_lab <- paste0("Y", seq(1, k))
  dat[time_lab] <- 0
  dat[, time_lab[1]] <- 1
  for (j in seq_along(time_lab)) {
    dat[time_k == j, time_lab[j]] <- 1
  }
  #names(dat)[names(dat)=="Y1"] <- "intercept"

  ## Select time knots and generate time mesh
  t_knots = seq(1, k, length = k)
  mesh1 = inla.mesh.1d(loc = t_knots, degree = 1, boundary = "free")

@@

    mesh2d <- INLA::inla.mesh.2d(
      offset = offset,
      boundary = bnd,
      max.edge = max_edge,
      cutoff = cutoff
    )
  }
  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], "), n = ", mesh$n)

  if (plot) {
    graphics::plot(mesh, asp = 1)
    graphics::points(coords, col = "red")
  }

@@

  # make projection points stack:
  sdat <- INLA::inla.stack(tag = "stdata",
                           data = list(y = dat$y),
                           A = A.list,
                           effects = effect.list
  )
# 
#   formula <- as.formula(paste(
#     "y ~ -1 + ",
#     paste(covar_names, collapse = "+"), 
#     "+ f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model = "ar1"))"))
#

  #covariates <- paste(covar_names, collapse = "+")

  formula <- y ~ -1 + depth_scaled+depth_scaled2+Y1+Y2+Y3+Y4+Y5+Y6 +
   f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model = "ar1"))

 # formula <- as.formula(paste(
 #    "y ~ -1 + ",
 #    paste(covar_names, collapse = "+"), "+ 
 #      f(year_f, model = "rw2") +
 #      f(year_f2, depth_scaled, model = "rw2") +
 #      f(year_f3, depth_scaled2, model = "rw2") +
 #      f(spatial.field, model=spde)"
 #  ))
 #

@@

```{r}
obj <- fit_inla_sensor(temp_dat, response = "temperature_C_avg",
                       n_knots = 50,
                       family = "gaussian",
                       plot = TRUE,
                       fit_model = TRUE,
                       extend = list(n = 8, offset = -0.1),
                       RangeGuess = 100, 
                       covariates = c("depth_scaled", "depth_scaled2"),
                       verbose = FALSE,
                       debug = FALSE,
                       trials = 2)

@@

```{r , eval=TRUE, include=FLASE}
# alternative to function, hard code
# samples = 10L
# pred_grid <- pg_one_year

>>>>

```{r}
testdat <- tidy_survey_sets(raw, c("SYN HS","SYN QCS","SYN WCVI","SYN WCHG"), years = c(2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018))
```

Get sensor data
```{r}
# get_ssids()
# sensor_syn <- get_sensor_data_trawl(ssid = c(1, 3, 4, 16), spread_attributes = TRUE)
# View(sensor_syn)
# saveRDS(sensor_syn, "data-cache/allsynoptic_sensor_07mar19_time2.rds")
# write.csv(sensor, "synoptic_3_sensor_data_07mar19.csv")
# sensor <- read_csv("data-cache/synoptic_3_sensor_data_07mar19.csv", col_types = cols(do_mlpL_avg = col_double(), 
#                     do_mlpL_max = col_double(), do_mlpL_min = col_double(), salinity_PSU_avg = col_double(), salinity_PSU_max = col_double(), 
#                     salinity_PSU_min = col_double()), na = "NA")
sensor <- readRDS("data-cache/allsynoptic_sensor_07mar19_time2.rds")
glimpse(sensor)
# View(sensor)
```

Join sensor data with trawl data from tidy_survey_sets function
```{r}
dat <- left_join(testdat, sensor, by="fishing_event_id", suffix = c("", ".sensor"))
dat$year_f <- factor(dat$year)
# View(dat)
dat <- dat %>% mutate(do_range = do_mlpL_max - do_mlpL_min, salinity_range = salinity_PSU_max -
  salinity_PSU_min, temp_range = temperature_C_max - temperature_C_min, depth_range = depth_m_max - depth_m_min)
# glimpse(dat)
```

Optional: Interpolate missing depth data
```{r, eval=FALSE}
testdat2 <- gfplot:::interp_survey_bathymetry(testdat)
glimpse(testdat2[[1]])
# if bathymetry ran...
dat <- left_join(testdat2[[1]], sensor, by="fishing_event_id", suffix = c("", ".sensor"))
dat$year_f <- factor(dat$year)
# View(dat)
dat <- dat %>% mutate(do_range = do_mlpL_max - do_mlpL_min, salinity_range = salinity_PSU_max -
  salinity_PSU_min, temp_range = temperature_C_max - temperature_C_min, depth_range = depth_m_max - depth_m_min)
```

@@

```{r}
scaled <- dat %>% filter(depth != "NA") %>% scale_survey_predictors()
# glimpse(scaled)
temp_dat <- scaled %>% filter (density!="NA") %>% filter ( ssid =="1") %>% filter(temperature_C_avg!="NA") # %>% filter(year==2017)
# glimpse(temp_dat)
# do_dat <- scaled %>% filter (density!="NA") %>% filter ( ssid =="1") %>%filter(do_mlpL_avg!="NA")
#pg <- readRDS("prediction-grids/QCSpred-grid-interp-cell-width-2.rds")
#pg <- as.array(pg)

@@

```{r}
ggplot(temp_dat, aes(depth)) + geom_histogram() 
ggplot(temp_dat, aes(density*1000)) + geom_histogram() 
ggplot(temp_dat, aes(depth, (density*1000))) + geom_point() 
#ggplot(synhs, aes(depth_m, log(density*1000+1))) + geom_point() 
```

@@

d <- temp_dat
response = "temperature_C_avg"
n_knots = 50
family = "gaussian"
plot = TRUE
fit_model = TRUE
extend = list(n = 8, offset = -0.1)
RangeGuess = 100 
offset = c(RangeGuess/5, RangeGuess)
max_edge = c(RangeGuess/5,RangeGuess)
cutoff = (RangeGuess/5)/5
prior_range = c(RangeGuess, 0.01) # P(range < 0.05) = 0.01
prior_sigma = c(0.5, 0.5)
kmeans = FALSE    
verbose = FALSE
debug = FALSE
trials = 2
covariates = c("depth_scaled", "depth_scaled2")
time = "year"
re_model_type = "rw1"
formula = full_model

fit_inla_sensor <- function(d, response = "temperature_C_avg",
                     n_knots = 50,
                     family = "binomial",
                     plot = FALSE,
                     fit_model = TRUE,
                     re_model_type = "rw1",
                     extend = list(n = 8, offset = -0.1),
                     kmeans = FALSE,                   
                     RangeGuess = 100, 
                     offset = c(RangeGuess/5, RangeGuess),                 
                     max_edge = c(RangeGuess/5,RangeGuess),
                     cutoff = (RangeGuess/5)/5,
                     prior_range = c(RangeGuess, 0.01), # P(range < 0.05) = 0.01
                     prior_sigma = c(0.5, 0.5),
                     time = "year",
                     covariates = NULL,
                     verbose = FALSE,
                     debug = FALSE,
                     trials = 10,
                     formula = full_model) {

  # create data.frame including response ("y"), coordinates("xcoo", "ycoo"), covariates, but not "intercept" as that will be input as time 1
  dat <- data.frame(
    y = d[, response, drop = TRUE],
    xcoo = d$X,
    ycoo = d$Y, 
    raw_time = d[, time, drop = TRUE],
    select(d, covariates)
  )

  # make sure dat is ordered temporally
  YO  <- order(dat$raw_time)
  dat <- dat[YO,]

  # number years sequentially, removing placeholders for missing years
  time_k <- as.numeric(as.factor(dat$raw_time))
  k <- max(time_k)

  # make design matrix with first time as intercept
  time_lab <- paste0("Y", seq(1, k)) # make vector of each year labeled "Y1", "Y2", ..."
  dat[time_lab] <- 0 # fill empty matrix with 0
  dat[, time_lab[1]] <- 1 # first time point assigned 1 for intercept
  for (j in seq_along(time_lab)) {
    dat[time_k == j, time_lab[j]] <- 1
  }

  ## Generate time mesh
  t_knots = seq(1, k, length = k)
  mesh1 = inla.mesh.1d(loc = t_knots, degree = 1, boundary = "free")

@@

    mesh2d <- INLA::inla.mesh.2d(
      offset = offset,
      boundary = bnd,
      max.edge = max_edge,
      cutoff = cutoff
    )
  }
  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], "), n = ", mesh2d$n)

  if (plot) {
    graphics::plot(mesh2d, asp = 1)
    graphics::points(coords, col = "red")
  }

@@

  # make projection points stack:
  sdat <- INLA::inla.stack(tag = "stdata",
                           data = list(y = dat$y),
                           A = A.list,
                           effects = effect.list
  )

  full_model <- as.formula(paste(
    'y ~ -1 + ',
    paste(covar_names, collapse = '+'),
    '+ f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model ="', re_model_type,'"))'))

@@

```{r}
obj <- fit_inla_sensor(temp_dat, response = "temperature_C_avg",
                       n_knots = 50,
                       family = "gaussian",
                       plot = TRUE,
                       fit_model = TRUE,
                       re_model_type = "rw1",
                       extend = list(n = 8, offset = -0.1),
                       RangeGuess = 100, 
                       covariates = c("depth_scaled", "depth_scaled2"),
                       verbose = FALSE,
                       debug = FALSE,
                       trials = 2
                       )

@@

```{r , eval=TRUE, include=FALSE}
# alternative to function, hard code
# samples = 10L
# pred_grid <- pg_one_year
